---
title: "A UN Report on the Relationship between GDP and Life Expectancy"
author: "Brianna Zichettella"
date: "2023-10-13"
output:
  word_document: default
  html_document: default
---


```{r setup, include = FALSE}
library(tidyverse)
library(here)
library(knitr)

options(scipen = 1000000)
#added because things kept printing in scientific notation
```

This report was prepared to the attention of the UN. It analyzes the relationship between a country's GDP, its life expectancy, and CO2 emissions. Our goal is to determine to what degree a country's economic strength or weakness is realated to its public heatlh status and impact on climate pollution. We hypothesize that obth life expectancy and CO2 emissions increase with the country's GDP. 


```{r data, include = FALSE}
gapminder_1997 <- read_csv(here("data/gapminder_1997.csv"))
```

```{r gdp_lifeexp_1997, echo = FALSE}
#echo means don't show the code but do run the code
ggplot(data = gapminder_1997, 
       aes(x = gdpPercap, y = lifeExp, 
           color = continent, 
           size = pop/1000000))+
  geom_point()+
  labs(x = "GDP per Capita", 
       y = "Life Expectancy", 
       title = "Do People in Wealthy Countries Live Longer?", 
       size = "Population (in millions)")
```

```{r, include=FALSE}
n_countries <- gapminder_1997 %>%
  select(country) %>% 
  n_distinct()
#n_distinct counts then number of values in the column specified by select

minimum_gdp <- gapminder_1997 %>% 
  summarize(min(gdpPercap)) %>% 
  round(digits = 1)

maximum_gdp <- gapminder_1997 %>% 
  summarize(max(gdpPercap)) %>%
  round(digits = 1)

#format - will format things nicely (we're not getting into that now though)
```

The above plot shows the relationship between GDP per capita and life expectancy for a total of `r n_countries` countries. For this set of countries, economic wealth ranged from a minimum of $`r minimum_gdp` to a maximum of \$`r maximum_gdp` per capita.

```{r table notes, include=FALSE}
#the backslash is there because rmarkdown puts formulas between dollar signs, so the backslash interrupts that but doesn't show up in markdown

#for table below, make something a header by putting dashes (?) under it
# can put something in column next to "summary of data" for a second header
#putting a colon on either side of the dashes (within the vertical lines) justifies the column, putting them on either side centers it
```

| Summary of data | |
|:----------|---------:|
|Number of Countries |`r n_countries`|
|Minimum GDP per Capita |`r minimum_gdp`|
|Maximum GDP per Capita | `r maximum_gdp`|

```{r table, echo=FALSE}
gapminder_1997 %>% 
  filter(continent == "Oceania") %>% 
  kable()
#kable is from the knitr library
```

Exercise 1: Headers and Sub-Headers

# Header
## Subheader

Exercise 2: 

Unnumbered list

* Programming language: **R**
* Function: *group_by*
* What's next? ***multiple imputation***
* Guide: R Markdown cheat sheet <https://www.rstudio.org/links/r_markdown_cheat_sheet>


Numbered list

1. **R**
2. *piping*
3. ***multiple imputation***
4. Behold: [R Markdown cheat sheet](https://www.rstudio.org/links/r_markdown_cheat_sheet)
5. You can also make all of the numbered entries 1's


